---
title: "CellChat"
output: html_notebook
---

load library
```{r message=FALSE, warning=FALSE}
library(Seurat)
library(SingleCellExperiment)
library(scran)
library(SeuratData)
library(edgeR)
library(limma)
library(clusterProfiler)
library(data.table)
library(SingleCellExperiment)
library(SummarizedExperiment)
library(muscat)
library(pROC)
library(igraph)
library(mclust)
library(cluster)
library(SCINA)
library(caret)
library(MAST)
library(AUCell)
library(HGNChelper)
library(SeuratDisk)
library(monocle3)
library(ggpubr)
library(slingshot)
library(grDevices)
library(RColorBrewer)
library(randomForest)
library(caTools)
library(mltools)
library(stringr)
library(reldist)
library(aricode)
library(TSCAN)
library(reshape2)
library(pheatmap)
library(philentropy)
library(rawr)
library(splatter)
library(combinat)
library(Hmisc)
library(CellChat)
```

read and write out as txt
```{r}
dat <- readRDS("test_CellChat/Mariana_Heart_DNMT3A_WT_FlorianLeuschner_25.08.22.Rds")

dat_out <- as.matrix(dat@assays$RNA@counts)
write.table(dat_out,"rawcount_txt/CellChat_DNMT3A.txt",quote=FALSE,sep="\t")
```

# test CellChat(raw lognorm)
read in
```{r}
dat <- readRDS("test_CellChat/Mariana_Heart_DNMT3A_WT_FlorianLeuschner_25.08.22.Rds")
dat <- NormalizeData(dat)
dat$samples <- as.factor(dat$sample)

CHIP <- subset(dat,condition=="DNMT3A")
NoCHIP <- subset(dat,condition=="WT")

data.input <- CHIP@assays$RNA@data # normalized data matrix
meta <- CHIP@meta.data
cellchatCHIP <- createCellChat(object=data.input,meta=meta,group.by="celltypes")

data.input <- NoCHIP@assays$RNA@data # normalized data matrix
meta <- NoCHIP@meta.data
cellchatNoCHIP <- createCellChat(object=data.input,meta=meta,group.by="celltypes")

CellChatDB <- CellChatDB.mouse # use CellChatDB.mouse if running on mouse data
# use all CellChatDB except for "Non-protein Signaling" for cell-cell communication analysis
CellChatDB.use <- subsetDB(CellChatDB)
cellchatCHIP@DB <- CellChatDB.use
cellchatNoCHIP@DB <- CellChatDB.use
```

pre-process
```{r}
cellchatCHIP <- subsetData(cellchatCHIP) # This step is necessary even if using the whole database
cellchatNoCHIP <- subsetData(cellchatNoCHIP) # This step is necessary even if using the whole database
future::plan("multisession",workers=4) # do parallel
cellchatCHIP <- identifyOverExpressedGenes(cellchatCHIP)
cellchatCHIP <- identifyOverExpressedInteractions(cellchatCHIP)
cellchatNoCHIP <- identifyOverExpressedGenes(cellchatNoCHIP)
cellchatNoCHIP <- identifyOverExpressedInteractions(cellchatNoCHIP)
```

get results(~30-40min)
```{r}
cellchatCHIP <- computeCommunProb(cellchatCHIP,type="triMean")
cellchatNoCHIP <- computeCommunProb(cellchatNoCHIP,type="triMean")
#df.net <- subsetCommunication(cellchatCHIP,sources.use=c("Monocytes","FB"),targets.use=c("Monocytes","FB"))
cellchatCHIP <- computeCommunProbPathway(cellchatCHIP)
cellchatNoCHIP <- computeCommunProbPathway(cellchatNoCHIP)
cellchatCHIP <- aggregateNet(cellchatCHIP)
cellchatNoCHIP <- aggregateNet(cellchatNoCHIP)

saveRDS(cellchatCHIP,"test_CellChat/DNMT3A_CHIP_Raw.rds")
saveRDS(cellchatNoCHIP,"test_CellChat/DNMT3A_NoCHIP_Raw.rds")
```

all int plots
```{r}
groupSize <- as.numeric(table(cellchatCHIP@idents))
par(mfrow = c(1,2), xpd=TRUE)
netVisual_circle(cellchatCHIP@net$count,vertex.weight=groupSize,weight.scale=T,label.edge=F,title.name="Number of interactions")
netVisual_circle(cellchatCHIP@net$weight,vertex.weight=groupSize,weight.scale=T,label.edge=F,title.name="Interaction weights/strength")
```


results and plots in original paper
network plot
```{r}
cellchatCHIP <- readRDS("test_CellChat/DNMT3A_CHIP_Raw.rds")
cellchatNoCHIP <- readRDS("test_CellChat/DNMT3A_NoCHIP_Raw.rds")

pathways.show <- c("EGF") 

#netVisual_aggregate(cellchatNoCHIP,signaling=pathways.show,sources.use=c("Monocytes"))
# Circle plot
netVisual_aggregate(cellchatNoCHIP,signaling=pathways.show,layout="circle",sources.use=c("Monocytes"))

#netVisual_aggregate(cellchatCHIP,signaling=pathways.show,sources.use=c("Monocytes"))
# Circle plot
netVisual_aggregate(cellchatCHIP,signaling=pathways.show,layout="circle",sources.use=c("Monocytes"))
```

bubble plot
```{r}
pairLR.use <- data.frame(interaction_name=c("NAMPT_ITGA5_ITGB1","RETN_TLR4","IL1B_IL1R1_IL1RAP","CCL3_CCR5","CCL3_CCR1","CCL4_CCR5","HBEGF_ERBB2_ERBB4","HBEGF_ERBB4","HBEGF_EGFR_ERBB2","HBEGF_EGFR","AREG_EGFR_ERBB2","AREG_EGFR"))
netVisual_bubble(cellchatNoCHIP,sources.use=c("Monocytes"),targets.use=c("FB"),pairLR.use=pairLR.use,thresh=1)
netVisual_bubble(cellchatCHIP,sources.use=c("Monocytes"),targets.use=c("FB"),pairLR.use=pairLR.use,thresh=1)
```

ranknet
```{r}
object.list <- list(NO_CHIP=cellchatNoCHIP,CHIP=cellchatCHIP)
cellchat <- mergeCellChat(object.list,add.names=names(object.list),cell.prefix=TRUE)

gg1 <- rankNet(cellchat,mode="comparison",sources.use="Monocytes",targets.use="FB",stacked=T,do.stat=FALSE,thresh=0.05,cutoff.pvalue=0.05,signaling=c('NPR2','IL1','EGF','RESISTIN','CCL','IGF','PARS','COMPLEMENT','SEMA4A'))
gg2 <- rankNet(cellchat,mode="comparison",sources.use="Monocytes",targets.use="FB",stacked=F,do.stat=FALSE,thresh=0.05,cutoff.pvalue=0.05,signaling=c('NPR2','IL1','EGF','RESISTIN','CCL','IGF','PARS','COMPLEMENT','SEMA4A'))
gg1+gg2
```

total interactions and p values
```{r}
cellchatNoCHIP@net$count
cellchatNoCHIP@net$weight
sum(cellchatNoCHIP@net$pval<0.05)
```


# test CellChat(afMF)
read in
```{r}
dat <- readRDS("test_CellChat/Mariana_Heart_DNMT3A_WT_FlorianLeuschner_25.08.22.Rds")
dat <- NormalizeData(dat)
dat$samples <- as.factor(dat$sample)

CHIP <- subset(dat,condition=="DNMT3A")
NoCHIP <- subset(dat,condition=="WT")

data.input <- CHIP@assays$RNA@data # normalized data matrix
meta <- CHIP@meta.data
cellchatCHIP <- createCellChat(object=data.input,meta=meta,group.by="celltypes")

data.input <- NoCHIP@assays$RNA@data # normalized data matrix
meta <- NoCHIP@meta.data
cellchatNoCHIP <- createCellChat(object=data.input,meta=meta,group.by="celltypes")

CellChatDB <- CellChatDB.mouse # use CellChatDB.mouse if running on mouse data
# use all CellChatDB except for "Non-protein Signaling" for cell-cell communication analysis
CellChatDB.use <- subsetDB(CellChatDB)
cellchatCHIP@DB <- CellChatDB.use
cellchatNoCHIP@DB <- CellChatDB.use
```

pre-process
```{r}
cellchatCHIP <- subsetData(cellchatCHIP) # This step is necessary even if using the whole database
cellchatNoCHIP <- subsetData(cellchatNoCHIP) # This step is necessary even if using the whole database
future::plan("multisession",workers=4) # do parallel
cellchatCHIP <- identifyOverExpressedGenes(cellchatCHIP)
cellchatCHIP <- identifyOverExpressedInteractions(cellchatCHIP)
cellchatNoCHIP <- identifyOverExpressedGenes(cellchatNoCHIP)
cellchatNoCHIP <- identifyOverExpressedInteractions(cellchatNoCHIP)
```

get results
```{r}
cellchatCHIP <- computeCommunProb(cellchatCHIP,type="triMean")
cellchatNoCHIP <- computeCommunProb(cellchatNoCHIP,type="triMean")
#df.net <- subsetCommunication(cellchatCHIP,sources.use=c("Monocytes","FB"),targets.use=c("Monocytes","FB"))
cellchatCHIP <- computeCommunProbPathway(cellchatCHIP)
cellchatNoCHIP <- computeCommunProbPathway(cellchatNoCHIP)
cellchatCHIP <- aggregateNet(cellchatCHIP)
cellchatNoCHIP <- aggregateNet(cellchatNoCHIP)

saveRDS(cellchatCHIP,"test_CellChat/DNMT3A_CHIP_Raw.rds")
saveRDS(cellchatNoCHIP,"test_CellChat/DNMT3A_NoCHIP_Raw.rds")
```

all int plots
```{r}
groupSize <- as.numeric(table(cellchatCHIP@idents))
par(mfrow = c(1,2), xpd=TRUE)
netVisual_circle(cellchatCHIP@net$count,vertex.weight=groupSize,weight.scale=T,label.edge=F,title.name="Number of interactions")
netVisual_circle(cellchatCHIP@net$weight,vertex.weight=groupSize,weight.scale=T,label.edge=F,title.name="Interaction weights/strength")
```


results and plots in original paper
network plot
```{r}
cellchatCHIP_afMF <- readRDS("test_CellChat/DNMT3A_CHIP_afMF.rds")
cellchatNoCHIP_afMF <- readRDS("test_CellChat/DNMT3A_NoCHIP_afMF.rds")

pathways.show <- c("EGF") 

#netVisual_aggregate(cellchatNoCHIP,signaling=pathways.show,sources.use=c("Monocytes"))
# Circle plot
netVisual_aggregate(cellchatNoCHIP_afMF,signaling=pathways.show,layout="circle",sources.use=c("Monocytes"))

#netVisual_aggregate(cellchatCHIP,signaling=pathways.show,sources.use=c("Monocytes"))
# Circle plot
netVisual_aggregate(cellchatCHIP_afMF,signaling=pathways.show,layout="circle",sources.use=c("Monocytes"))

```

bubble plot
```{r fig.height=2, fig.width=1.5}
pairLR.use <- data.frame(interaction_name=c("NAMPT_ITGA5_ITGB1","RETN_TLR4","IL1B_IL1R1_IL1RAP","CCL3_CCR5","CCL3_CCR1","CCL4_CCR5","HBEGF_ERBB2_ERBB4","HBEGF_ERBB4","HBEGF_EGFR_ERBB2","HBEGF_EGFR","AREG_EGFR_ERBB2","AREG_EGFR"))
netVisual_bubble(cellchatNoCHIP_afMF,sources.use=c("Monocytes"),targets.use=c("FB"),pairLR.use=pairLR.use,thresh=1)
netVisual_bubble(cellchatCHIP_afMF,sources.use=c("Monocytes"),targets.use=c("FB"),pairLR.use=pairLR.use,thresh=1)
#netVisual_bubble(cellchatCHIP,sources.use=c("Monocytes"),targets.use=c("FB"))
```

ranknet
```{r}
object.list <- list(NO_CHIP=cellchatNoCHIP_afMF,CHIP=cellchatCHIP_afMF)
cellchat_afMF <- mergeCellChat(object.list,add.names=names(object.list),cell.prefix=TRUE)

gg1 <- rankNet(cellchat_afMF,mode="comparison",sources.use="Monocytes",targets.use="FB",stacked=T,do.stat=FALSE,thresh=0.05,cutoff.pvalue=0.05,signaling=c('NPR2','IL1','EGF','RESISTIN','CCL','IGF','PARS','COMPLEMENT','SEMA4A'))
gg2 <- rankNet(cellchat_afMF,mode="comparison",sources.use="Monocytes",targets.use="FB",stacked=F,do.stat=FALSE,thresh=0.05,cutoff.pvalue=0.05,signaling=c('NPR2','IL1','EGF','RESISTIN','CCL','IGF','PARS','COMPLEMENT','SEMA4A'))
gg1+gg2
```


total interactions and p values
```{r}
cellchatNoCHIP_afMF@net$count
cellchatNoCHIP_afMF@net$weight
sum(cellchatNoCHIP_afMF@net$pval<0.05)
```


# test CellChat(ALRA)
read in
```{r}
dat <- readRDS("test_CellChat/Mariana_Heart_DNMT3A_WT_FlorianLeuschner_25.08.22.Rds")
dat <- NormalizeData(dat)
dat$samples <- as.factor(dat$sample)

CHIP <- subset(dat,condition=="DNMT3A")
NoCHIP <- subset(dat,condition=="WT")

data.input <- CHIP@assays$RNA@data # normalized data matrix
meta <- CHIP@meta.data
cellchatCHIP <- createCellChat(object=data.input,meta=meta,group.by="celltypes")

data.input <- NoCHIP@assays$RNA@data # normalized data matrix
meta <- NoCHIP@meta.data
cellchatNoCHIP <- createCellChat(object=data.input,meta=meta,group.by="celltypes")

CellChatDB <- CellChatDB.mouse # use CellChatDB.mouse if running on mouse data
# use all CellChatDB except for "Non-protein Signaling" for cell-cell communication analysis
CellChatDB.use <- subsetDB(CellChatDB)
cellchatCHIP@DB <- CellChatDB.use
cellchatNoCHIP@DB <- CellChatDB.use
```

pre-process
```{r}
cellchatCHIP <- subsetData(cellchatCHIP) # This step is necessary even if using the whole database
cellchatNoCHIP <- subsetData(cellchatNoCHIP) # This step is necessary even if using the whole database
future::plan("multisession",workers=4) # do parallel
cellchatCHIP <- identifyOverExpressedGenes(cellchatCHIP)
cellchatCHIP <- identifyOverExpressedInteractions(cellchatCHIP)
cellchatNoCHIP <- identifyOverExpressedGenes(cellchatNoCHIP)
cellchatNoCHIP <- identifyOverExpressedInteractions(cellchatNoCHIP)
```

get results
```{r}
cellchatCHIP <- computeCommunProb(cellchatCHIP,type="triMean")
cellchatNoCHIP <- computeCommunProb(cellchatNoCHIP,type="triMean")
#df.net <- subsetCommunication(cellchatCHIP,sources.use=c("Monocytes","FB"),targets.use=c("Monocytes","FB"))
cellchatCHIP <- computeCommunProbPathway(cellchatCHIP)
cellchatNoCHIP <- computeCommunProbPathway(cellchatNoCHIP)
cellchatCHIP <- aggregateNet(cellchatCHIP)
cellchatNoCHIP <- aggregateNet(cellchatNoCHIP)

saveRDS(cellchatCHIP,"test_CellChat/DNMT3A_CHIP_Raw.rds")
saveRDS(cellchatNoCHIP,"test_CellChat/DNMT3A_NoCHIP_Raw.rds")
```

all int plots
```{r}
groupSize <- as.numeric(table(cellchatCHIP@idents))
par(mfrow = c(1,2), xpd=TRUE)
netVisual_circle(cellchatCHIP@net$count,vertex.weight=groupSize,weight.scale=T,label.edge=F,title.name="Number of interactions")
netVisual_circle(cellchatCHIP@net$weight,vertex.weight=groupSize,weight.scale=T,label.edge=F,title.name="Interaction weights/strength")
```


results and plots in original paper
network plot
```{r}
cellchatCHIP_ALRA <- readRDS("test_CellChat/DNMT3A_CHIP_ALRA.rds")
cellchatNoCHIP_ALRA <- readRDS("test_CellChat/DNMT3A_NoCHIP_ALRA.rds")

pathways.show <- c("EGF") 

#netVisual_aggregate(cellchatNoCHIP,signaling=pathways.show,sources.use=c("Monocytes"))
# Circle plot
netVisual_aggregate(cellchatNoCHIP_ALRA,signaling=pathways.show,layout="circle",sources.use=c("Monocytes"))

#netVisual_aggregate(cellchatCHIP,signaling=pathways.show,sources.use=c("Monocytes"))
# Circle plot
netVisual_aggregate(cellchatCHIP_ALRA,signaling=pathways.show,layout="circle",sources.use=c("Monocytes"))

```

bubble plot
```{r fig.height=2, fig.width=1.5}
pairLR.use <- data.frame(interaction_name=c("NAMPT_ITGA5_ITGB1","RETN_TLR4","IL1B_IL1R1_IL1RAP","CCL3_CCR5","CCL3_CCR1","CCL4_CCR5","HBEGF_ERBB2_ERBB4","HBEGF_ERBB4","HBEGF_EGFR_ERBB2","HBEGF_EGFR","AREG_EGFR_ERBB2","AREG_EGFR"))
netVisual_bubble(cellchatNoCHIP_ALRA,sources.use=c("Monocytes"),targets.use=c("FB"),pairLR.use=pairLR.use,thresh=1)
netVisual_bubble(cellchatCHIP_ALRA,sources.use=c("Monocytes"),targets.use=c("FB"),pairLR.use=pairLR.use,thresh=1)
#netVisual_bubble(cellchatCHIP,sources.use=c("Monocytes"),targets.use=c("FB"))
```

ranknet
```{r}
object.list <- list(NO_CHIP=cellchatNoCHIP_afMF,CHIP=cellchatCHIP_afMF)
cellchat_afMF <- mergeCellChat(object.list,add.names=names(object.list),cell.prefix=TRUE)

gg1 <- rankNet(cellchat_afMF,mode="comparison",sources.use="Monocytes",targets.use="FB",stacked=T,do.stat=FALSE,thresh=0.05,cutoff.pvalue=0.05,signaling=c('NPR2','IL1','EGF','RESISTIN','CCL','IGF','PARS','COMPLEMENT','SEMA4A'))
gg2 <- rankNet(cellchat_afMF,mode="comparison",sources.use="Monocytes",targets.use="FB",stacked=F,do.stat=FALSE,thresh=0.05,cutoff.pvalue=0.05,signaling=c('NPR2','IL1','EGF','RESISTIN','CCL','IGF','PARS','COMPLEMENT','SEMA4A'))
gg1+gg2
```


total interactions and p values
```{r}
cellchatNoCHIP_afMF@net$count
cellchatNoCHIP_afMF@net$weight
sum(cellchatNoCHIP_afMF@net$pval<0.05)
```

# test CellChat(MAGIC)
read in
```{r}
dat <- readRDS("test_CellChat/Mariana_Heart_DNMT3A_WT_FlorianLeuschner_25.08.22.Rds")
dat$samples <- as.factor(dat$sample)

imputed_dat <- fread("imputed/alra/CellChat_DNMT3A_alra.txt",sep="\t",header=TRUE,stringsAsFactors=FALSE)
imputed_dat <- as.matrix(imputed_dat,rownames=1)
colnames(imputed_dat) <- colnames(dat)

dat <- CreateSeuratObject(imputed_dat,meta.data=dat@meta.data,min.cells=0,min.features=0)

CHIP <- subset(dat,condition=="DNMT3A")
NoCHIP <- subset(dat,condition=="WT")

data.input <- CHIP@assays$RNA@counts # already normalized data matrix
meta <- CHIP@meta.data
cellchatCHIP <- createCellChat(object=data.input,meta=meta,group.by="celltypes")

data.input <- NoCHIP@assays$RNA@counts # already normalized data matrix
meta <- NoCHIP@meta.data
cellchatNoCHIP <- createCellChat(object=data.input,meta=meta,group.by="celltypes")

CellChatDB <- CellChatDB.mouse # use CellChatDB.mouse if running on mouse data
# use all CellChatDB except for "Non-protein Signaling" for cell-cell communication analysis
CellChatDB.use <- subsetDB(CellChatDB)
cellchatCHIP@DB <- CellChatDB.use
cellchatNoCHIP@DB <- CellChatDB.use
```

pre-process
```{r}
cellchatCHIP <- subsetData(cellchatCHIP) # This step is necessary even if using the whole database
cellchatNoCHIP <- subsetData(cellchatNoCHIP) # This step is necessary even if using the whole database
future::plan("multisession",workers=4) # do parallel
cellchatCHIP <- identifyOverExpressedGenes(cellchatCHIP)
cellchatCHIP <- identifyOverExpressedInteractions(cellchatCHIP)
cellchatNoCHIP <- identifyOverExpressedGenes(cellchatNoCHIP)
cellchatNoCHIP <- identifyOverExpressedInteractions(cellchatNoCHIP)
```

get results(batch job)
```{r}
cellchatCHIP <- computeCommunProb(cellchatCHIP,type="triMean")
cellchatNoCHIP <- computeCommunProb(cellchatNoCHIP,type="triMean")
#df.net <- subsetCommunication(cellchatCHIP,sources.use=c("Monocytes","FB"),targets.use=c("Monocytes","FB"))
cellchatCHIP <- computeCommunProbPathway(cellchatCHIP)
cellchatNoCHIP <- computeCommunProbPathway(cellchatNoCHIP)
cellchatCHIP <- aggregateNet(cellchatCHIP)
cellchatNoCHIP <- aggregateNet(cellchatNoCHIP)

saveRDS(cellchatCHIP,"test_CellChat/DNMT3A_CHIP_ALRA.rds")
saveRDS(cellchatNoCHIP,"test_CellChat/DNMT3A_NoCHIP_ALRA.rds")
```

all int plots
```{r}
groupSize <- as.numeric(table(cellchatCHIP@idents))
par(mfrow = c(1,2), xpd=TRUE)
netVisual_circle(cellchatCHIP@net$count,vertex.weight=groupSize,weight.scale=T,label.edge=F,title.name="Number of interactions")
netVisual_circle(cellchatCHIP@net$weight,vertex.weight=groupSize,weight.scale=T,label.edge=F,title.name="Interaction weights/strength")
```


results and plots in original paper
network plot
```{r}
cellchatCHIP <- readRDS("test_CellChat/DNMT3A_CHIP_MAGIC.rds")
cellchatNoCHIP <- readRDS("test_CellChat/DNMT3A_NoCHIP_MAGIC.rds")

pathways.show <- c("EGF") 

#netVisual_aggregate(cellchatNoCHIP,signaling=pathways.show,idents.use=c("Monocytes","FB"))
# Circle plot
netVisual_aggregate(cellchatNoCHIP,signaling=pathways.show,layout="circle",sources.use=c("Monocytes"))

#netVisual_aggregate(cellchatCHIP,signaling=pathways.show,idents.use=c("Monocytes","FB"))
# Circle plot
netVisual_aggregate(cellchatCHIP,signaling=pathways.show,layout="circle",sources.use=c("Monocytes"))
```

bubble plot
```{r}
pairLR.use <- data.frame(interaction_name=c("NAMPT_ITGA5_ITGB1","RETN_TLR4","IL1B_IL1R1_IL1RAP","CCL3_CCR5","CCL3_CCR1","CCL4_CCR5","HBEGF_ERBB2_ERBB4","HBEGF_ERBB4","HBEGF_EGFR_ERBB2","HBEGF_EGFR","AREG_EGFR_ERBB2","AREG_EGFR"))
netVisual_bubble(cellchatCHIP,sources.use=c("Monocytes"),targets.use=c("FB"),pairLR.use=pairLR.use,thresh=1)
netVisual_bubble(cellchatNoCHIP,sources.use=c("Monocytes"),targets.use=c("FB"),pairLR.use=pairLR.use,thresh=1)
#netVisual_bubble(cellchatCHIP,sources.use=c("Monocytes"),targets.use=c("FB"))
```

# test CellChat(AutoClass)
read in
```{r}
dat <- readRDS("test_CellChat/Mariana_Heart_DNMT3A_WT_FlorianLeuschner_25.08.22.Rds")
dat$samples <- as.factor(dat$sample)

imputed_dat <- fread("imputed/alra/CellChat_DNMT3A_alra.txt",sep="\t",header=TRUE,stringsAsFactors=FALSE)
imputed_dat <- as.matrix(imputed_dat,rownames=1)
colnames(imputed_dat) <- colnames(dat)

dat <- CreateSeuratObject(imputed_dat,meta.data=dat@meta.data,min.cells=0,min.features=0)

CHIP <- subset(dat,condition=="DNMT3A")
NoCHIP <- subset(dat,condition=="WT")

data.input <- CHIP@assays$RNA@counts # already normalized data matrix
meta <- CHIP@meta.data
cellchatCHIP <- createCellChat(object=data.input,meta=meta,group.by="celltypes")

data.input <- NoCHIP@assays$RNA@counts # already normalized data matrix
meta <- NoCHIP@meta.data
cellchatNoCHIP <- createCellChat(object=data.input,meta=meta,group.by="celltypes")

CellChatDB <- CellChatDB.mouse # use CellChatDB.mouse if running on mouse data
# use all CellChatDB except for "Non-protein Signaling" for cell-cell communication analysis
CellChatDB.use <- subsetDB(CellChatDB)
cellchatCHIP@DB <- CellChatDB.use
cellchatNoCHIP@DB <- CellChatDB.use
```

pre-process
```{r}
cellchatCHIP <- subsetData(cellchatCHIP) # This step is necessary even if using the whole database
cellchatNoCHIP <- subsetData(cellchatNoCHIP) # This step is necessary even if using the whole database
future::plan("multisession",workers=4) # do parallel
cellchatCHIP <- identifyOverExpressedGenes(cellchatCHIP)
cellchatCHIP <- identifyOverExpressedInteractions(cellchatCHIP)
cellchatNoCHIP <- identifyOverExpressedGenes(cellchatNoCHIP)
cellchatNoCHIP <- identifyOverExpressedInteractions(cellchatNoCHIP)
```

get results(use batch job)
```{r}
cellchatCHIP <- computeCommunProb(cellchatCHIP,type="triMean")
cellchatNoCHIP <- computeCommunProb(cellchatNoCHIP,type="triMean")
#df.net <- subsetCommunication(cellchatCHIP,sources.use=c("Monocytes","FB"),targets.use=c("Monocytes","FB"))
cellchatCHIP <- computeCommunProbPathway(cellchatCHIP)
cellchatNoCHIP <- computeCommunProbPathway(cellchatNoCHIP)
cellchatCHIP <- aggregateNet(cellchatCHIP)
cellchatNoCHIP <- aggregateNet(cellchatNoCHIP)

saveRDS(cellchatCHIP,"test_CellChat/DNMT3A_CHIP_ALRA.rds")
saveRDS(cellchatNoCHIP,"test_CellChat/DNMT3A_NoCHIP_ALRA.rds")
```

all int plots
```{r}
groupSize <- as.numeric(table(cellchatCHIP@idents))
par(mfrow = c(1,2), xpd=TRUE)
netVisual_circle(cellchatCHIP@net$count,vertex.weight=groupSize,weight.scale=T,label.edge=F,title.name="Number of interactions")
netVisual_circle(cellchatCHIP@net$weight,vertex.weight=groupSize,weight.scale=T,label.edge=F,title.name="Interaction weights/strength")
```


results and plots in original paper
network plot
```{r}
cellchatCHIP <- readRDS("test_CellChat/DNMT3A_CHIP_AC.rds")
cellchatNoCHIP <- readRDS("test_CellChat/DNMT3A_NoCHIP_AC.rds")

pathways.show <- c("EGF") 

#netVisual_aggregate(cellchatNoCHIP,signaling=pathways.show,idents.use=c("Monocytes","FB"))
# Circle plot
netVisual_aggregate(cellchatNoCHIP,signaling=pathways.show,layout="circle",sources.use=c("Monocytes"))

#netVisual_aggregate(cellchatCHIP,signaling=pathways.show,idents.use=c("Monocytes","FB"))
# Circle plot
netVisual_aggregate(cellchatCHIP,signaling=pathways.show,layout="circle",sources.use=c("Monocytes"))


#netVisual_aggregate(cellchatNoCHIP,signaling=pathways.show)
#netVisual_aggregate(cellchatCHIP,signaling=pathways.show)
```

bubble plot
```{r}
pairLR.use <- data.frame(interaction_name=c("NAMPT_ITGA5_ITGB1","RETN_TLR4","IL1B_IL1R1_IL1RAP","CCL3_CCR5","CCL3_CCR1","CCL4_CCR5","HBEGF_ERBB2_ERBB4","HBEGF_ERBB4","HBEGF_EGFR_ERBB2","HBEGF_EGFR","AREG_EGFR_ERBB2","AREG_EGFR"))
netVisual_bubble(cellchatCHIP,sources.use=c("Monocytes"),targets.use=c("FB"),pairLR.use=pairLR.use,thresh=1)
netVisual_bubble(cellchatNoCHIP,sources.use=c("Monocytes"),targets.use=c("FB"),pairLR.use=pairLR.use,thresh=1)
#netVisual_bubble(cellchatCHIP,sources.use=c("Monocytes"),targets.use=c("FB"))
```


# Analysis
read data
```{r}
cellchatCHIP <- readRDS("test_CellChat/DNMT3A_CHIP_Raw.rds")
cellchatNoCHIP <- readRDS("test_CellChat/DNMT3A_NoCHIP_Raw.rds")
object.list <- list(NO_CHIP=cellchatNoCHIP,CHIP=cellchatCHIP)
cellchat <- mergeCellChat(object.list,add.names=names(object.list),cell.prefix=TRUE)

cellchatCHIP_afMF <- readRDS("test_CellChat/DNMT3A_CHIP_afMF.rds")
cellchatNoCHIP_afMF <- readRDS("test_CellChat/DNMT3A_NoCHIP_afMF.rds")
object.list <- list(NO_CHIP=cellchatNoCHIP_afMF,CHIP=cellchatCHIP_afMF)
cellchat_afMF <- mergeCellChat(object.list,add.names=names(object.list),cell.prefix=TRUE)

cellchatCHIP_ALRA <- readRDS("test_CellChat/DNMT3A_CHIP_ALRA.rds")
cellchatNoCHIP_ALRA <- readRDS("test_CellChat/DNMT3A_NoCHIP_ALRA.rds")
object.list <- list(NO_CHIP=cellchatNoCHIP_ALRA,CHIP=cellchatCHIP_ALRA)
cellchat_ALRA <- mergeCellChat(object.list,add.names=names(object.list),cell.prefix=TRUE)

cellchatCHIP_MAGIC <- readRDS("test_CellChat/DNMT3A_CHIP_MAGIC.rds")
cellchatNoCHIP_MAGIC <- readRDS("test_CellChat/DNMT3A_NoCHIP_MAGIC.rds")
object.list <- list(NO_CHIP=cellchatNoCHIP_MAGIC,CHIP=cellchatCHIP_MAGIC)
cellchat_MAGIC <- mergeCellChat(object.list,add.names=names(object.list),cell.prefix=TRUE)

cellchatCHIP_MAGIClog <- readRDS("test_CellChat/DNMT3A_CHIP_MAGIClog.rds")
cellchatNoCHIP_MAGIClog <- readRDS("test_CellChat/DNMT3A_NoCHIP_MAGIClog.rds")
object.list <- list(NO_CHIP=cellchatNoCHIP_MAGIClog,CHIP=cellchatCHIP_MAGIClog)
cellchat_MAGIClog <- mergeCellChat(object.list,add.names=names(object.list),cell.prefix=TRUE)

cellchatCHIP_AC <- readRDS("test_CellChat/DNMT3A_CHIP_AC.rds")
cellchatNoCHIP_AC <- readRDS("test_CellChat/DNMT3A_NoCHIP_AC.rds")
object.list <- list(NO_CHIP=cellchatNoCHIP_AC,CHIP=cellchatCHIP_AC)
cellchat_AC <- mergeCellChat(object.list,add.names=names(object.list),cell.prefix=TRUE)
```

network
```{r}
pathways.show <- c("EGF") 

p1 <- netVisual_aggregate(cellchatNoCHIP,signaling=pathways.show,layout="circle",sources.use=c("Monocytes"))
p2 <- netVisual_aggregate(cellchatCHIP,signaling=pathways.show,layout="circle",sources.use=c("Monocytes"))
p3 <- netVisual_aggregate(cellchatNoCHIP_afMF,signaling=pathways.show,layout="circle",sources.use=c("Monocytes"))
p4 <- netVisual_aggregate(cellchatCHIP_afMF,signaling=pathways.show,layout="circle",sources.use=c("Monocytes"))
p5 <- netVisual_aggregate(cellchatNoCHIP_ALRA,signaling=pathways.show,layout="circle",sources.use=c("Monocytes"))
p6 <- netVisual_aggregate(cellchatCHIP_ALRA,signaling=pathways.show,layout="circle",sources.use=c("Monocytes"))
p7 <- netVisual_aggregate(cellchatNoCHIP_AC,signaling=pathways.show,layout="circle",sources.use=c("Monocytes"))
p8 <- netVisual_aggregate(cellchatCHIP_AC,signaling=pathways.show,layout="circle",sources.use=c("Monocytes"))
p9 <- netVisual_aggregate(cellchatNoCHIP_MAGIC,signaling=pathways.show,layout="circle",sources.use=c("Monocytes"))
p10 <- netVisual_aggregate(cellchatCHIP_MAGIC,signaling=pathways.show,layout="circle",sources.use=c("Monocytes"))
p11 <- netVisual_aggregate(cellchatNoCHIP_MAGIClog,signaling=pathways.show,layout="circle",sources.use=c("Monocytes"))
p12 <- netVisual_aggregate(cellchatCHIP_MAGIClog,signaling=pathways.show,layout="circle",sources.use=c("Monocytes"))


```

bubble plot
```{r fig.width=2.8}
pairLR.use <- data.frame(interaction_name=c("NAMPT_ITGA5_ITGB1","RETN_TLR4","IL1B_IL1R1_IL1RAP","CCL3_CCR5","CCL3_CCR1","CCL4_CCR5","HBEGF_ERBB2_ERBB4","HBEGF_ERBB4","HBEGF_EGFR_ERBB2","HBEGF_EGFR","AREG_EGFR_ERBB2","AREG_EGFR"))

p1 <- netVisual_bubble(cellchatNoCHIP,sources.use=c("Monocytes"),targets.use=c("FB"),pairLR.use=pairLR.use,thresh=1)
p2 <- netVisual_bubble(cellchatCHIP,sources.use=c("Monocytes"),targets.use=c("FB"),pairLR.use=pairLR.use,thresh=1)
p3 <- netVisual_bubble(cellchatNoCHIP_afMF,sources.use=c("Monocytes"),targets.use=c("FB"),pairLR.use=pairLR.use,thresh=1)
p4 <- netVisual_bubble(cellchatCHIP_afMF,sources.use=c("Monocytes"),targets.use=c("FB"),pairLR.use=pairLR.use,thresh=1)
p5 <- netVisual_bubble(cellchatNoCHIP_ALRA,sources.use=c("Monocytes"),targets.use=c("FB"),pairLR.use=pairLR.use,thresh=1)
p6 <- netVisual_bubble(cellchatCHIP_ALRA,sources.use=c("Monocytes"),targets.use=c("FB"),pairLR.use=pairLR.use,thresh=1)
p7 <- netVisual_bubble(cellchatNoCHIP_AC,sources.use=c("Monocytes"),targets.use=c("FB"),pairLR.use=pairLR.use,thresh=1)
p8 <- netVisual_bubble(cellchatCHIP_AC,sources.use=c("Monocytes"),targets.use=c("FB"),pairLR.use=pairLR.use,thresh=1)
p9 <- netVisual_bubble(cellchatNoCHIP_MAGIC,sources.use=c("Monocytes"),targets.use=c("FB"),pairLR.use=pairLR.use,thresh=1)
p10 <- netVisual_bubble(cellchatCHIP_MAGIC,sources.use=c("Monocytes"),targets.use=c("FB"),pairLR.use=pairLR.use,thresh=1)
p11 <- netVisual_bubble(cellchatNoCHIP_MAGIClog,sources.use=c("Monocytes"),targets.use=c("FB"),pairLR.use=pairLR.use,thresh=1)
p12 <- netVisual_bubble(cellchatCHIP_MAGIClog,sources.use=c("Monocytes"),targets.use=c("FB"),pairLR.use=pairLR.use,thresh=1)
p1|p2
p3|p4
p5|p6
p7|p8
p9|p10
p11|p12
```

ranknet
```{r fig.width=5}
gg1 <- rankNet(cellchat,mode="comparison",sources.use="Monocytes",targets.use="FB",stacked=T,do.stat=FALSE,thresh=0.05,cutoff.pvalue=0.05,signaling=c('NPR2','IL1','EGF','RESISTIN','CCL','IGF','PARS','COMPLEMENT','SEMA4A'))
gg2 <- rankNet(cellchat,mode="comparison",sources.use="Monocytes",targets.use="FB",stacked=F,do.stat=FALSE,thresh=0.05,cutoff.pvalue=0.05,signaling=c('NPR2','IL1','EGF','RESISTIN','CCL','IGF','PARS','COMPLEMENT','SEMA4A'))
gg1+gg2

gg3 <- rankNet(cellchat_afMF,mode="comparison",sources.use="Monocytes",targets.use="FB",stacked=T,do.stat=FALSE,thresh=0.05,cutoff.pvalue=0.05,signaling=c('NPR2','IL1','EGF','RESISTIN','CCL','IGF','PARS','COMPLEMENT','SEMA4A'))
gg4 <- rankNet(cellchat_afMF,mode="comparison",sources.use="Monocytes",targets.use="FB",stacked=F,do.stat=FALSE,thresh=0.05,cutoff.pvalue=0.05,signaling=c('NPR2','IL1','EGF','RESISTIN','CCL','IGF','PARS','COMPLEMENT','SEMA4A'))
gg3+gg4

gg5 <- rankNet(cellchat_ALRA,mode="comparison",sources.use="Monocytes",targets.use="FB",stacked=T,do.stat=FALSE,thresh=0.05,cutoff.pvalue=0.05,signaling=c('NPR2','IL1','EGF','RESISTIN','CCL','IGF','PARS','COMPLEMENT','SEMA4A'))
gg6 <- rankNet(cellchat_ALRA,mode="comparison",sources.use="Monocytes",targets.use="FB",stacked=F,do.stat=FALSE,thresh=0.05,cutoff.pvalue=0.05,signaling=c('NPR2','IL1','EGF','RESISTIN','CCL','IGF','PARS','COMPLEMENT','SEMA4A'))
gg5+gg6

gg7 <- rankNet(cellchat_AC,mode="comparison",sources.use="Monocytes",targets.use="FB",stacked=T,do.stat=FALSE,thresh=0.05,cutoff.pvalue=0.05,signaling=c('NPR2','IL1','EGF','RESISTIN','CCL','IGF','PARS','COMPLEMENT','SEMA4A'))
gg8 <- rankNet(cellchat_AC,mode="comparison",sources.use="Monocytes",targets.use="FB",stacked=F,do.stat=FALSE,thresh=0.05,cutoff.pvalue=0.05,signaling=c('NPR2','IL1','EGF','RESISTIN','CCL','IGF','PARS','COMPLEMENT','SEMA4A'))
gg7+gg8

gg9 <- rankNet(cellchat_MAGIC,mode="comparison",sources.use="Monocytes",targets.use="FB",stacked=T,do.stat=FALSE,thresh=0.05,cutoff.pvalue=0.05,signaling=c('NPR2','IL1','EGF','RESISTIN','CCL','IGF','PARS','COMPLEMENT','SEMA4A'))
gg10 <- rankNet(cellchat_MAGIC,mode="comparison",sources.use="Monocytes",targets.use="FB",stacked=F,do.stat=FALSE,thresh=0.05,cutoff.pvalue=0.05,signaling=c('NPR2','IL1','EGF','RESISTIN','CCL','IGF','PARS','COMPLEMENT','SEMA4A'))
gg9+gg10

gg11 <- rankNet(cellchat_MAGIClog,mode="comparison",sources.use="Monocytes",targets.use="FB",stacked=T,do.stat=FALSE,thresh=0.05,cutoff.pvalue=0.05,signaling=c('NPR2','IL1','EGF','RESISTIN','CCL','IGF','PARS','COMPLEMENT','SEMA4A'))
gg12 <- rankNet(cellchat_MAGIClog,mode="comparison",sources.use="Monocytes",targets.use="FB",stacked=F,do.stat=FALSE,thresh=0.05,cutoff.pvalue=0.05,signaling=c('NPR2','IL1','EGF','RESISTIN','CCL','IGF','PARS','COMPLEMENT','SEMA4A'))
gg11+gg12
```

numbers of interactions
```{r fig.height=4.5, fig.width=2.5}
#all
df_NumInt <- data.frame(Method=rep(c("Raw","afMF","ALRA","AutoClass","MAGIC","MAGIC_log"),each=2),Type=rep(c("NO_CHIP","CHIP"),6),Number=c(sum(cellchatNoCHIP@net$pval<0.05),sum(cellchatCHIP@net$pval<0.05),sum(cellchatNoCHIP_afMF@net$pval<0.05),sum(cellchatCHIP_afMF@net$pval<0.05),sum(cellchatNoCHIP_ALRA@net$pval<0.05),sum(cellchatCHIP_ALRA@net$pval<0.05),sum(cellchatNoCHIP_AC@net$pval<0.05),sum(cellchatCHIP_AC@net$pval<0.05),sum(cellchatNoCHIP_MAGIC@net$pval<0.05),sum(cellchatCHIP_MAGIC@net$pval<0.05),sum(cellchatNoCHIP_MAGIClog@net$pval<0.05),sum(cellchatCHIP_MAGIClog@net$pval<0.05)))

NumInt_all <- ggbarplot(df_NumInt,x="Method",y="Number",fill="Type",palette="npg",position=position_dodge(0.9))+theme(axis.text.x=element_text(angle=90))+xlab("Method")+ylab("Interactions(P<0.05),n")+ggtitle("All Interactions(P<0.05)")#+coord_cartesian(ylim=c(0.6,1))

#monocytes
df_NumInt <- data.frame(Method=rep(c("Raw","afMF","ALRA","AutoClass","MAGIC","MAGIC_log"),each=2),Type=rep(c("NO_CHIP","CHIP"),6),Number=c(sum(cellchatNoCHIP@net$pval['Monocytes',,]<0.05),sum(cellchatCHIP@net$pval['Monocytes',,]<0.05),sum(cellchatNoCHIP_afMF@net$pval['Monocytes',,]<0.05),sum(cellchatCHIP_afMF@net$pval['Monocytes',,]<0.05),sum(cellchatNoCHIP_ALRA@net$pval['Monocytes',,]<0.05),sum(cellchatCHIP_ALRA@net$pval['Monocytes',,]<0.05),sum(cellchatNoCHIP_AC@net$pval['Monocytes',,]<0.05),sum(cellchatCHIP_AC@net$pval['Monocytes',,]<0.05),sum(cellchatNoCHIP_MAGIC@net$pval['Monocytes',,]<0.05),sum(cellchatCHIP_MAGIC@net$pval['Monocytes',,]<0.05),sum(cellchatNoCHIP_MAGIClog@net$pval['Monocytes',,]<0.05),sum(cellchatCHIP_MAGIClog@net$pval['Monocytes',,]<0.05)))

NumInt_mono <- ggbarplot(df_NumInt,x="Method",y="Number",fill="Type",palette="npg",position=position_dodge(0.9))+theme(axis.text.x=element_text(angle=90))+xlab("Method")+ylab("Interactions(P<0.05),n")+ggtitle("From-Monocyte Interactions(P<0.05)")#+coord_cartesian(ylim=c(0.6,1))

##percentages
df_NumInt <- data.frame(Method=rep(c("Raw","afMF","ALRA","AutoClass","MAGIC","MAGIC_log"),each=2),Type=rep(c("NO_CHIP","CHIP"),6),PCT=c(sum(cellchatNoCHIP@net$pval<0.05)/(dim(cellchatNoCHIP@net$pval)[1]*dim(cellchatNoCHIP@net$pval)[2]*dim(cellchatNoCHIP@net$pval)[3]),sum(cellchatCHIP@net$pval<0.05)/(dim(cellchatCHIP@net$pval)[1]*dim(cellchatCHIP@net$pval)[2]*dim(cellchatCHIP@net$pval)[3]),sum(cellchatNoCHIP_afMF@net$pval<0.05)/(dim(cellchatNoCHIP_afMF@net$pval)[1]*dim(cellchatNoCHIP_afMF@net$pval)[2]*dim(cellchatNoCHIP_afMF@net$pval)[3]),sum(cellchatCHIP_afMF@net$pval<0.05)/(dim(cellchatCHIP_afMF@net$pval)[1]*dim(cellchatCHIP_afMF@net$pval)[2]*dim(cellchatCHIP_afMF@net$pval)[3]),sum(cellchatNoCHIP_ALRA@net$pval<0.05)/(dim(cellchatNoCHIP_ALRA@net$pval)[1]*dim(cellchatNoCHIP_ALRA@net$pval)[2]*dim(cellchatNoCHIP_ALRA@net$pval)[3]),sum(cellchatCHIP_ALRA@net$pval<0.05)/(dim(cellchatCHIP_ALRA@net$pval)[1]*dim(cellchatCHIP_ALRA@net$pval)[2]*dim(cellchatCHIP_ALRA@net$pval)[3]),sum(cellchatNoCHIP_AC@net$pval<0.05)/(dim(cellchatNoCHIP_AC@net$pval)[1]*dim(cellchatNoCHIP_AC@net$pval)[2]*dim(cellchatNoCHIP_AC@net$pval)[3]),sum(cellchatCHIP_AC@net$pval<0.05)/(dim(cellchatCHIP_AC@net$pval)[1]*dim(cellchatCHIP_AC@net$pval)[2]*dim(cellchatCHIP_AC@net$pval)[3]),sum(cellchatNoCHIP_MAGIC@net$pval<0.05)/(dim(cellchatNoCHIP_MAGIC@net$pval)[1]*dim(cellchatNoCHIP_MAGIC@net$pval)[2]*dim(cellchatNoCHIP_MAGIC@net$pval)[3]),sum(cellchatCHIP_MAGIC@net$pval<0.05)/(dim(cellchatCHIP_MAGIC@net$pval)[1]*dim(cellchatCHIP_MAGIC@net$pval)[2]*dim(cellchatCHIP_MAGIC@net$pval)[3]),sum(cellchatNoCHIP_MAGIClog@net$pval<0.05)/(dim(cellchatNoCHIP_MAGIClog@net$pval)[1]*dim(cellchatNoCHIP_MAGIClog@net$pval)[2]*dim(cellchatNoCHIP_MAGIClog@net$pval)[3]),sum(cellchatCHIP_MAGIClog@net$pval<0.05)/(dim(cellchatCHIP_MAGIClog@net$pval)[1]*dim(cellchatCHIP_MAGIClog@net$pval)[2]*dim(cellchatCHIP_MAGIClog@net$pval)[3])))

Pct_all <- ggbarplot(df_NumInt,x="Method",y="PCT",fill="Type",palette="npg",position=position_dodge(0.9))+theme(axis.text.x=element_text(angle=90))+xlab("Method")+ylab("Interactions(P<0.05),%")+ggtitle("% of All P<0.05 Interactions")#+coord_cartesian(ylim=c(0.6,1))

ggarrange(NumInt_all,Pct_all,NumInt_mono,ncol=1,nrow=3,labels="AUTO")
```



