---
title: "Clustering_Cellcycle"
output: html_notebook
---

load library
```{r message=FALSE, warning=FALSE}
library(Seurat)
library(ggplot2)
library(SingleCellExperiment)
library(scran)
library(SeuratData)
library(edgeR)
library(limma)
library(clusterProfiler)
library(data.table)
library(SingleCellExperiment)
library(SummarizedExperiment)
library(muscat)
library(pROC)
library(igraph)
library(mclust)
library(cluster)
library(SCINA)
library(caret)
library(robCompositions)
library(easyCODA)
library(MAST)
library(AUCell)
library(HGNChelper)
library(SeuratDisk)
library(scmap)
library(monocle3)
library(ggpubr)
library(slingshot)
library(grDevices)
library(RColorBrewer)
library(randomForest)
library(caTools)
library(mltools)
#library(stringr)
library(reldist)
library(aricode)
library(TSCAN)
library(reshape2)
library(pheatmap)
library(philentropy)
library(SCeQTL)
library(eQTLsingle)
library(rawr)
library(splatter)
library(combinat)

selected_palette <- c("aquamarine4","azure4","powderblue","burlywood","coral","cornflowerblue","cyan3","darkgoldenrod2","darkorange3","peachpuff","royalblue","khaki2","plum1","lightsalmon")
selected_palette2 <- c("plum1","aquamarine4","azure4","powderblue","burlywood","coral","cornflowerblue","cyan3","darkgoldenrod2","darkorange3","peachpuff","royalblue")
```

clustering functions
```{r message=FALSE, warning=FALSE}
getCluster <- function(x,meta,reso=0.01,npc=10,cent=4){
    tmp <- CreateSeuratObject(counts=x,meta.data=meta,min.cells=0,min.features=0)
    tmp <- FindVariableFeatures(tmp,nfeatures=3000,verbose=FALSE)
    tmp <- ScaleData(tmp,verbose=FALSE)
    tmp <- RunPCA(tmp,verbose=FALSE)
    tmp$kmeans <- kmeans(tmp@reductions$pca@cell.embeddings[,1:npc],centers=cent)$cluster
    tmp <- FindNeighbors(tmp,dims=1:npc,verbose=FALSE)
    tmp <- FindClusters(tmp,resolution=reso,verbose=FALSE)
    tmpgraph <- buildSNNGraph(tmp@reductions$pca@cell.embeddings[,1:npc],transposed=T,k=10,d=NA)
    res <- cluster_louvain(tmpgraph)$membership
    cc <- aggregate(tmp@reductions$pca@cell.embeddings[,1:npc],list(res),mean)
    cc <- as.matrix(cc[,-1])
    hclu <- hclust(dist(cc))
    clu <- cutree(hclu,cent)
    clu <- clu[res]      
    tmp$lv_clu <- clu
    #tmp <- tmp@meta.data
    tmp <- RunUMAP(tmp,dims=1:npc,verbose=FALSE)
    return(tmp)
}

evalClustering <- function(x,npc=10,method="lv_clu"){
    
    acc <- -mean(sapply(unique(x@meta.data[,method]),function(i){
        p <- table(x@meta.data[x@meta.data[,method]==i,"celltype"])/sum(x@meta.data[,method]==i)
        sum(p*log(p),na.rm=TRUE)
    }))

    pur <- -mean(sapply(unique(x$celltype),function(sct){
        p <- table(x@meta.data[x$celltype==sct,method])/sum(x$celltype==sct)
        sum(p*log(p),na.rm=TRUE)
    }))

    ARI <- adjustedRandIndex(x$celltype,x@meta.data[,method])

    d <- dist(x@reductions$pca@cell.embeddings[,1:npc])
    s <- silhouette(x@meta.data[,method],dist=d)
    medianSil <- median(s[,3])
    
    NMI_value <- NMI(x@meta.data[,method],x$celltype)
    #NMI_value <- NMI(x$lv_clu,x$celltype,variant="sum")
    
    df <- c(acc,pur,ARI,medianSil,NMI_value)
    names(df) <- c("acc","pur","ARI","medianSil","NMI")
    return(df)
}

evalClustering_status <- function(x,npc=10,method="lv_clu"){
    
    acc <- -mean(sapply(unique(x@meta.data[,method]),function(i){
        p <- table(x@meta.data[x@meta.data[,method]==i,"label"])/sum(x@meta.data[,method]==i)
        sum(p*log(p),na.rm=TRUE)
    }))

    pur <- -mean(sapply(unique(x$label),function(sct){
        p <- table(x@meta.data[x$label==sct,method])/sum(x$label==sct)
        sum(p*log(p),na.rm=TRUE)
    }))

    ARI <- adjustedRandIndex(x$label,x@meta.data[,method])

    d <- dist(x@reductions$pca@cell.embeddings[,1:npc])
    s <- silhouette(x@meta.data[,method],dist=d)
    medianSil <- median(s[,3])
    
    NMI_value <- NMI(x@meta.data[,method],x$label)
    #NMI_value <- NMI(x$lv_clu,x$celltype,variant="sum")
    
    df <- c(acc,pur,ARI,medianSil,NMI_value)
    names(df) <- c("acc","pur","ARI","medianSil","NMI")
    return(df)
}
```

# clustering for mixture samples
## GSE75748 cell type
cell level
read data
```{r message=FALSE, warning=FALSE}
##bulk
bulk_dat <- read.csv("bulk/GSE75748_bulk_cell_type_ec.csv",header=TRUE,row.names=1)
bulk_dat <- DGEList(bulk_dat)
keep <- filterByExpr(bulk_dat)
bulk_dat <- bulk_dat[keep,,keep.lib.sizes=FALSE]
bulk_dat <- calcNormFactors(bulk_dat)
bulk_meta <- c("H1","H1","H1","H1","H9","H9","H9","DEC","DEC","EC","EC","EC","HFF","HFF","HFF","NPC","NPC","TB","TB") 
mm <- model.matrix(~0 + bulk_meta)
bulk_dat <- voom(bulk_dat,mm,plot=FALSE)$E
bulk_dat <- data.frame(H1=apply(bulk_dat[,1:4],1,mean),H9=apply(bulk_dat[,5:7],1,mean),DEC=apply(bulk_dat[,8:9],1,mean),EC=apply(bulk_dat[,10:12],1,mean),HFF=apply(bulk_dat[,13:15],1,mean),NPC=apply(bulk_dat[,16:17],1,mean),TB=apply(bulk_dat[,18:19],1,mean))

##raw
raw_dat <- fread("rawcount_txt/GSE75748_sc_cell_type_genebycell.txt",sep="\t",header=TRUE,stringsAsFactors=FALSE)
raw_dat <- as.matrix(raw_dat,rownames=1)
##imputed
MAGIC_dat <- fread("MAGIC/GSE75748_sc_cell_type_genebycell.tsv",sep="\t",header=TRUE,stringsAsFactors=FALSE)
MAGIC_dat <- as.matrix(MAGIC_dat,rownames=1)
alra_dat <- fread("alra/GSE75748_sc_cell_type_genebycell_alra.txt",sep="\t",stringsAsFactors=FALSE)
alra_dat <- as.matrix(alra_dat,rownames=1)
dca_dat <- fread("dca/GSE75748_sc_cell_type/mean.tsv",sep="\t",header=TRUE,stringsAsFactors=FALSE)
dca_dat <- as.matrix(dca_dat,rownames=1)
MAGIClog_dat <- fread("MAGIC_log/GSE75748_sc_cell_type_genebycell.tsv",sep="\t",header=TRUE,stringsAsFactors=FALSE)
MAGIClog_dat <- as.matrix(MAGIClog_dat,rownames=1)
AutoClass_dat <- fread("AutoClass/GSE75748_sc_cell_type_genebycell_AC.txt",sep="\t",header=TRUE,stringsAsFactors=FALSE)
AutoClass_dat <- as.matrix(AutoClass_dat,rownames=1)
ccImpute_dat <- fread("ccImpute/GSE75748_sc_cell_type_genebycell_cc.txt",sep="\t",header=TRUE,stringsAsFactors=FALSE)
ccImpute_dat <- as.matrix(ccImpute_dat,rownames=1)
ks_dat <- fread("knn_smoothing/GSE75748_sc_cell_type_genebycell_ks.txt",sep="\t",header=TRUE,stringsAsFactors=FALSE)
ks_dat <- as.matrix(ks_dat,rownames=1)
Iimpute_dat <- fread("I_Impute/GSE75748_sc_cell_type_genebycell_CImp.txt",sep="\t",header=TRUE,stringsAsFactors=FALSE)
Iimpute_dat <- as.matrix(Iimpute_dat,rownames=1)
scRMDnorm_dat <- fread("scRMD/GSE75748_sc_cell_type_genebycell_scRMDlognorm.txt",sep="\t",header=TRUE,stringsAsFactors=FALSE)
scRMDnorm_dat <- as.matrix(scRMDnorm_dat,rownames=1)
Bf_dat <- fread("Bfimpute/GSE75748_sc_cell_type_genebycell_Bf_specc.txt",sep="\t",header=TRUE,stringsAsFactors=FALSE)
Bf_dat <- as.matrix(Bf_dat,rownames=1)
afMF_dat <- fread("afMF/imputed_data/GSE75748_sc_cell_type_sigma3_0_convergence_True.txt",sep="\t",header=TRUE,stringsAsFactors=FALSE)
afMF_dat <- as.matrix(afMF_dat,rownames=1)

metadat <- sapply(colnames(raw_dat),function(x) strsplit(x,"\\_")[[1]][1])
metadat <- data.frame(cell=colnames(raw_dat),celltype=metadat)
```

normalized/log2 data
```{r}
raw_dat <- CreateSeuratObject(counts=raw_dat,meta.data=metadat,min.cells=0,min.features=0)
raw_dat <- NormalizeData(raw_dat)
raw_dat <- as.matrix(raw_dat@assays$RNA@data)

dca_dat <- log2(dca_dat)

ks_dat <- CreateSeuratObject(counts=ks_dat,meta.data=metadat,min.cells=0,min.features=0)
ks_dat <- NormalizeData(ks_dat)
ks_dat <- as.matrix(ks_dat@assays$RNA@data)

Iimpute_dat <- log2(Iimpute_dat)
```

combine normalized data
```{r}
dat <- list(raw=raw_dat,ALRA=alra_dat,AutoClass=AutoClass_dat,Bfimpute=Bf_dat,ccImpute=ccImpute_dat,DCA=dca_dat,I_Impute=Iimpute_dat,kNN_smoothing=ks_dat,MAGIC=MAGIC_dat,MAGIC_log=MAGIClog_dat,scRMD=scRMDnorm_dat,afMF=afMF_dat)
rm(raw_dat)
rm(alra_dat)
rm(AutoClass_dat)
rm(Bf_dat)
rm(ccImpute_dat)
rm(Iimpute_dat)
rm(ks_dat)
rm(dca_dat)
rm(MAGIC_dat)
rm(MAGIClog_dat)
rm(scRMDnorm_dat)
rm(afMF_dat)

#remove inconsistant genes
colnames(dat$ALRA) <- rownames(metadat)
bulk_dat <- bulk_dat[rownames(bulk_dat)%in%rownames(dat$DCA),]
dat <- lapply(dat,function(x) x[rownames(bulk_dat),] )
```

cluster
```{r message=FALSE, warning=FALSE}
datCluster <- lapply(dat,getCluster,metadat,cent=7)
```

evaluate lv
```{r}
res1_lv <- as.data.frame(lapply(datCluster,evalClustering))
as.data.frame(t(res1_lv))
```

evaluate km
```{r}
res1_km <- as.data.frame(lapply(datCluster,evalClustering,method="kmeans"))
as.data.frame(t(res1_km))
```

writeout/read
```{r}
#writeout
all_res <- list(lv=res1_lv,km=res1_km)
saveRDS(all_res,"results/SC_Clustering/SC_Clustering_EvalRes_GSE75748.rds")
```

UMAP lv celltype
```{r}
for(x in names(datCluster)){
    p <- DimPlot(datCluster[[x]],reduction="umap",group.by="celltype")+ggtitle(x)
    print(p)
}
```

PCA cell type
```{r}
for(x in names(datCluster)){
    p <- DimPlot(datCluster[[x]],reduction="pca",group.by="celltype")+ggtitle(x)
    print(p)
}
```

## GSE81861
read data
```{r message=FALSE, warning=FALSE}
bulk_dat <- readRDS("bulk/bulk_of_GSE81861_with_replicates_TPM.rds")
bulk_dat <- bulk_dat[rowSums(bulk_dat>0)>0,]
filterGene <- which(apply(bulk_dat,1,max)<5)
bulk_dat <- bulk_dat[-filterGene,]
bulk_meta <- c("K562","K562","A549","A549","GM12878","GM12878","K562","K562","IMR90","IMR90","H1","H1","GM12878","GM12878") 
bulk_dat <- log2(bulk_dat+1)

##raw
raw_dat <- fread("rawcount_txt/GSE81861_ENCODE_genebycell.txt",sep="\t",header=TRUE,stringsAsFactors=FALSE)
raw_dat <- as.matrix(raw_dat,rownames=1)
##imputed
MAGIC_dat <- fread("MAGIC/GSE81861_ENCODE_genebycell.tsv",sep="\t",header=TRUE,stringsAsFactors=FALSE)
MAGIC_dat <- as.matrix(MAGIC_dat,rownames=1)
alra_dat <- fread("alra/GSE81861_ENCODE_genebycell_alra.txt",sep="\t",stringsAsFactors=FALSE)
alra_dat <- as.matrix(alra_dat,rownames=1)
dca_dat <- fread("dca/GSE81861_ENCODE/mean.tsv",sep="\t",header=TRUE,stringsAsFactors=FALSE)
dca_dat <- as.matrix(dca_dat,rownames=1)
MAGIClog_dat <- fread("MAGIC_log/GSE81861_ENCODE_genebycell.tsv",sep="\t",header=TRUE,stringsAsFactors=FALSE)
MAGIClog_dat <- as.matrix(MAGIClog_dat,rownames=1)
AutoClass_dat <- fread("AutoClass/GSE81861_ENCODE_genebycell_AC.txt",sep="\t",header=TRUE,stringsAsFactors=FALSE)
AutoClass_dat <- as.matrix(AutoClass_dat,rownames=1)
ccImpute_dat <- fread("ccImpute/GSE81861_ENCODE_genebycell_cc.txt",sep="\t",header=TRUE,stringsAsFactors=FALSE)
ccImpute_dat <- as.matrix(ccImpute_dat,rownames=1)
ks_dat <- fread("knn_smoothing/GSE81861_ENCODE_genebycell_ks.txt",sep="\t",header=TRUE,stringsAsFactors=FALSE)
ks_dat <- as.matrix(ks_dat,rownames=1)
Iimpute_dat <- fread("I_Impute/GSE81861_ENCODE_genebycell_CImp.txt",sep="\t",header=TRUE,stringsAsFactors=FALSE)
Iimpute_dat <- as.matrix(Iimpute_dat,rownames=1)
scRMDnorm_dat <- fread("scRMD/GSE81861_ENCODE_genebycell_scRMDlognorm.txt",sep="\t",header=TRUE,stringsAsFactors=FALSE)
scRMDnorm_dat <- as.matrix(scRMDnorm_dat,rownames=1)
Bf_dat <- fread("Bfimpute/GSE81861_ENCODE_genebycell_Bf_specc.txt",sep="\t",header=TRUE,stringsAsFactors=FALSE)
Bf_dat <- as.matrix(Bf_dat,rownames=1)
afMF_dat <- fread("afMF/imputed_data/GSE81861_ENCODE_sigma3_0_convergence_True.txt",sep="\t",header=TRUE,stringsAsFactors=FALSE)
afMF_dat <- as.matrix(afMF_dat,rownames=1)

metadat <- sapply(colnames(raw_dat),function(x) strsplit(x,"\\_")[[1]][1])
metadat <- data.frame(cell=colnames(raw_dat),celltype=metadat)
```

normalized/log2 data
```{r}
raw_dat <- CreateSeuratObject(counts=raw_dat,meta.data=metadat,min.cells=0,min.features=0)
raw_dat <- NormalizeData(raw_dat)
raw_dat <- as.matrix(raw_dat@assays$RNA@data)

dca_dat <- log2(dca_dat)

ks_dat <- CreateSeuratObject(counts=ks_dat,meta.data=metadat,min.cells=0,min.features=0)
ks_dat <- NormalizeData(ks_dat)
ks_dat <- as.matrix(ks_dat@assays$RNA@data)

Iimpute_dat <- log2(Iimpute_dat)
```

combine normalized data
```{r}
dat <- list(raw=raw_dat,ALRA=alra_dat,AutoClass=AutoClass_dat,Bfimpute=Bf_dat,ccImpute=ccImpute_dat,DCA=dca_dat,I_Impute=Iimpute_dat,kNN_smoothing=ks_dat,MAGIC=MAGIC_dat,MAGIC_log=MAGIClog_dat,scRMD=scRMDnorm_dat,afMF=afMF_dat)
rm(raw_dat)
rm(alra_dat)
rm(AutoClass_dat)
rm(Bf_dat)
rm(ccImpute_dat)
rm(Iimpute_dat)
rm(ks_dat)
rm(dca_dat)
rm(MAGIC_dat)
rm(MAGIClog_dat)
rm(scRMDnorm_dat)
rm(afMF_dat)

#remove inconsistant genes
colnames(dat$ALRA) <- rownames(metadat)
bulk_dat <- bulk_dat[rownames(bulk_dat)%in%rownames(dat$DCA),]
dat <- lapply(dat,function(x) x[rownames(bulk_dat),] )
```

cluster
```{r message=FALSE, warning=FALSE}
datCluster <- lapply(dat,getCluster,metadat,cent=5)
```

evaluate lv
```{r}
res1_lv <- as.data.frame(lapply(datCluster,evalClustering))
as.data.frame(t(res1_lv))
```

evaluate km
```{r}
res1_km <- as.data.frame(lapply(datCluster,evalClustering,method="kmeans"))
as.data.frame(t(res1_km))
```

writeout/read
```{r}
#writeout
all_res <- list(lv=res1_lv,km=res1_km)
saveRDS(all_res,"results/SC_Clustering/SC_Clustering_EvalRes_GSE81861.rds")
```

UMAP lv celltype
```{r}
for(x in names(datCluster)){
    p <- DimPlot(datCluster[[x]],reduction="umap",group.by="celltype")+ggtitle(x)
    print(p)
}
```

PCA cell type
```{r}
for(x in names(datCluster)){
    p <- DimPlot(datCluster[[x]],reduction="pca",group.by="celltype")+ggtitle(x)
    print(p)
}
```

## cell bench
read data
```{r message=FALSE, warning=FALSE}
##bulk
bulk_dat <- readRDS("bulk/cellbench/GSE86337_processed_count.rds")
bulk_dat <- DGEList(bulk_dat)
bulk_dat <- calcNormFactors(bulk_dat)
bulk_meta <- c("HCC827","HCC827","H2228","H2228","H838","H838","A549","A549","H1975","H1975") 

##raw
raw_dat <- fread("rawcount_txt/cellbench_sc_10x_5cl_genebycell.txt",sep="\t",header=TRUE,stringsAsFactors=FALSE)
raw_dat <- as.matrix(raw_dat,rownames=1)
##imputed
MAGIC_dat <- fread("MAGIC/cellbench_sc_10x_5cl_genebycell.tsv",sep="\t",header=TRUE,stringsAsFactors=FALSE)
MAGIC_dat <- as.matrix(MAGIC_dat,rownames=1)
alra_dat <- fread("alra/cellbench_sc_10x_5cl_genebycell_alra.txt",sep="\t",stringsAsFactors=FALSE)
alra_dat <- as.matrix(alra_dat,rownames=1)
dca_dat <- fread("dca/cellbench_sc_10x_5cl/mean.tsv",sep="\t",header=TRUE,stringsAsFactors=FALSE)
dca_dat <- as.matrix(dca_dat,rownames=1)
MAGIClog_dat <- fread("MAGIC_log/cellbench_sc_10x_5cl_genebycell.tsv",sep="\t",header=TRUE,stringsAsFactors=FALSE)
MAGIClog_dat <- as.matrix(MAGIClog_dat,rownames=1)
AutoClass_dat <- fread("AutoClass/cellbench_sc_10x_5cl_genebycell_AC.txt",sep="\t",header=TRUE,stringsAsFactors=FALSE)
AutoClass_dat <- as.matrix(AutoClass_dat,rownames=1)
ccImpute_dat <- fread("ccImpute/cellbench_sc_10x_5cl_genebycell_cc.txt",sep="\t",header=TRUE,stringsAsFactors=FALSE)
ccImpute_dat <- as.matrix(ccImpute_dat,rownames=1)
ks_dat <- fread("knn_smoothing/cellbench_sc_10x_5cl_genebycell_ks.txt",sep="\t",header=TRUE,stringsAsFactors=FALSE)
ks_dat <- as.matrix(ks_dat,rownames=1)
Iimpute_dat <- fread("I_Impute/cellbench_sc_10x_5cl_genebycell_CImp.txt",sep="\t",header=TRUE,stringsAsFactors=FALSE)
Iimpute_dat <- as.matrix(Iimpute_dat,rownames=1)
scRMDnorm_dat <- fread("scRMD/cellbench_sc_10x_5cl_genebycell_scRMDlognorm.txt",sep="\t",header=TRUE,stringsAsFactors=FALSE)
scRMDnorm_dat <- as.matrix(scRMDnorm_dat,rownames=1)
Bf_dat <- fread("Bfimpute/cellbench_sc_10x_5cl_genebycell_Bf_specc.txt",sep="\t",header=TRUE,stringsAsFactors=FALSE)
Bf_dat <- as.matrix(Bf_dat,rownames=1)
afMF_dat <- fread("afMF/imputed_data/cellbench_sc_10x_5cl_sigma3_0_convergence_True.txt",sep="\t",header=TRUE,stringsAsFactors=FALSE)
afMF_dat <- as.matrix(afMF_dat,rownames=1)

metadat <- sapply(colnames(raw_dat),function(x) strsplit(x,"\\.")[[1]][2])
metadat <- data.frame(cell=colnames(raw_dat),celltype=metadat)
```

normalized/log2 data
```{r}
raw_dat <- CreateSeuratObject(counts=raw_dat,meta.data=metadat,min.cells=0,min.features=0)
raw_dat <- NormalizeData(raw_dat)
raw_dat <- as.matrix(raw_dat@assays$RNA@data)

dca_dat <- log2(dca_dat)

ks_dat <- CreateSeuratObject(counts=ks_dat,meta.data=metadat,min.cells=0,min.features=0)
ks_dat <- NormalizeData(ks_dat)
ks_dat <- as.matrix(ks_dat@assays$RNA@data)

Iimpute_dat <- log2(Iimpute_dat+min(Iimpute_dat[Iimpute_dat!=0]))
```

combine normalized data
```{r}
dat <- list(raw=raw_dat,ALRA=alra_dat,AutoClass=AutoClass_dat,Bfimpute=Bf_dat,ccImpute=ccImpute_dat,DCA=dca_dat,I_Impute=Iimpute_dat,kNN_smoothing=ks_dat,MAGIC=MAGIC_dat,MAGIC_log=MAGIClog_dat,scRMD=scRMDnorm_dat,afMF=afMF_dat)
rm(raw_dat)
rm(alra_dat)
rm(AutoClass_dat)
rm(Bf_dat)
rm(ccImpute_dat)
rm(Iimpute_dat)
rm(ks_dat)
rm(dca_dat)
rm(MAGIC_dat)
rm(MAGIClog_dat)
rm(scRMDnorm_dat)
rm(afMF_dat)

#remove inconsistant genes
colnames(dat$ALRA) <- rownames(metadat)
bulk_dat <- bulk_dat[rownames(bulk_dat)%in%rownames(dat$DCA),]
dat <- lapply(dat,function(x) x[rownames(bulk_dat),] )
```


cluster
```{r message=FALSE, warning=FALSE}
datCluster <- lapply(dat,getCluster,metadat,cent=5)
```

evaluate lv
```{r}
res1_lv <- as.data.frame(lapply(datCluster,evalClustering))
as.data.frame(t(res1_lv))
```

evaluate km
```{r}
res1_km <- as.data.frame(lapply(datCluster,evalClustering,method="kmeans"))
as.data.frame(t(res1_km))
```

writeout/read
```{r}
#writeout
all_res <- list(lv=res1_lv,km=res1_km)
saveRDS(all_res,"results/SC_Clustering/SC_Clustering_EvalRes_cellbench.rds")
```

UMAP celltype
```{r fig.height=10,fig.width=10}
p_list <- list()
for(x in names(datCluster)[names(datCluster)!="New"]){
    p_list[[x]] <- DimPlot(datCluster[[x]],reduction="umap",group.by="celltype")+ggtitle(x)+scale_color_brewer(palette="Set2")
    #print(p)
}
clu_cellbench_umap_ct <- ggarrange(plotlist=p_list[c("raw","AutoClass","DCA","afMF","ALRA","ccImpute","I_Impute","Bfimpute","scRMD","kNN_smoothing","MAGIC_log","MAGIC")],nrow=4,ncol=3)
clu_cellbench_umap_ct
```

UMAP lv clu
```{r fig.height=10,fig.width=10}
p_list <- list()
for(x in names(datCluster)[names(datCluster)!="New"]){
    p_list[[x]] <- DimPlot(datCluster[[x]],reduction="umap",group.by="lv_clu")+ggtitle(x)+scale_color_brewer(palette="Set2")
    #print(p)
}
clu_cellbench_umap_lv <- ggarrange(plotlist=p_list[c("raw","AutoClass","DCA","afMF","ALRA","ccImpute","I_Impute","Bfimpute","scRMD","kNN_smoothing","MAGIC_log","MAGIC")],nrow=4,ncol=3)
clu_cellbench_umap_lv
```

PCA cell type
```{r}
for(x in names(datCluster)){
    p <- DimPlot(datCluster[[x]],reduction="pca",group.by="celltype")+ggtitle(x)
    print(p)
}
```

## summary and visualization
read data
```{r}
d1 <- readRDS("results/SC_Clustering/SC_Clustering_EvalRes_GSE75748.rds")
d2 <- readRDS("results/SC_Clustering/SC_Clustering_EvalRes_GSE81861.rds")
d3 <- readRDS("results/SC_Clustering/SC_Clustering_EvalRes_cellbench.rds")

res1_lv <- d1$lv
res2_lv <- d2$lv
res3_lv <- d3$lv

res1_km <- d1$km
res2_km <- d2$km
res3_km <- d3$km
```

### process data lv
```{r}
##LV cluster
rownames(res1_lv) <- c("Hacc","Hpur","ARI","medianSil","NMI")

res_lv <- data.frame()
res_lv_median <- data.frame()
for(i in 1:2){
    tmp1 <- res1_lv[i,1]-res1_lv[i,-1]
    tmp2 <- res2_lv[i,1]-res2_lv[i,-1]
    tmp3 <- res3_lv[i,1]-res3_lv[i,-1]
    tmp <- rbind(tmp1,tmp2,tmp3)
    rownames(tmp) <- paste0(rownames(res1_lv)[i],"_",c("GSE75748","GSE81861","cellbench"))
    tmp_median <- apply(tmp,2,median)
    res_lv <- rbind(res_lv,tmp)
    res_lv_median <- rbind(res_lv_median,tmp_median)
}

for(i in 3:5){
    tmp1 <- res1_lv[i,-1]-res1_lv[i,1]
    tmp2 <- res2_lv[i,-1]-res2_lv[i,1]
    tmp3 <- res3_lv[i,-1]-res3_lv[i,1]
    tmp <- rbind(tmp1,tmp2,tmp3)
    rownames(tmp) <- paste0(rownames(res1_lv)[i],"_",c("GSE75748","GSE81861","cellbench"))
    tmp_median <- apply(tmp,2,median)
    res_lv <- rbind(res_lv,tmp)
    res_lv_median <- rbind(res_lv_median,tmp_median)
}

res_lv_median <- as.data.frame(t(res_lv_median))
rownames(res_lv_median) <- colnames(res1_lv)[-1]
colnames(res_lv_median) <- c("Hacc","Hpur","ARI","medianSil","NMI")

res_lv_median$method <- rownames(res_lv_median)
res_lv_median_melt <- data.table::melt(res_lv_median,id.vars=c("method"))
colnames(res_lv_median_melt)[2:3] <- c("Metrics","Metrics_Difference")
head(res_lv_median_melt)

res_lv$Metrics_Datasets <- rownames(res_lv)
res_lv$Metrics <- sapply(rownames(res_lv),function(x) strsplit(x,"\\_")[[1]][1] )
res_lv_melt <- data.table::melt(res_lv,id.vars=c("Metrics_Datasets","Metrics"))
colnames(res_lv_melt)[3:4] <- c("method","Metrics_Difference")
head(res_lv_melt)
```

heatmap(median metrics)
```{r}
rank1 <- apply(res_lv_median[,-ncol(res_lv_median)],2,rank)
rank1 <- apply(rank1,1,median)
rank1 <- names(rank1[order(rank1)])
res_lv_median_melt$Method <- factor(res_lv_median_melt$method,levels=unique(res_lv_median_melt$method[order(res_lv_median_melt$method)]))
res_lv_median_melt$method <- factor(res_lv_median_melt$method,levels=rank1)

tmp_hm <- res_lv_median_melt[,]
tmp_hm[tmp_hm$Metrics_Difference<(-0.03),]$Metrics_Difference <- -0.03
tmp_hm[tmp_hm$Metrics_Difference>(0.03),]$Metrics_Difference <- 0.03
tmp_hm <- tmp_hm[tmp_hm$Metrics!="medianSil",]
clu_4metric_lv_hm <- ggplot(tmp_hm,aes(method,Metrics,fill=Metrics_Difference))+geom_tile()+theme(axis.text.x=element_text(angle=90))+scale_fill_gradient2(low="blue",mid="white",high="red",midpoint=0)+ggtitle("Louvain Clustering Evaluation Metrics(Values of Difference)")+xlab("Method")+theme(legend.title=element_text(size=15),legend.text=element_text(size=14),axis.text=element_text(size=15),axis.title=element_text(size=15),plot.title=element_text(size=15))#+geom_vline(xintercept=neg_md+0.5,linetype="dashed")
clu_4metric_lv_hm
```

### process data kmeans
```{r}
##LV cluster
rownames(res1_km) <- c("Hacc","Hpur","ARI","medianSil","NMI")

res_km <- data.frame()
res_km_median <- data.frame()
for(i in 1:2){
    tmp1 <- res1_km[i,1]-res1_km[i,-1]
    tmp2 <- res2_km[i,1]-res2_km[i,-1]
    tmp3 <- res3_km[i,1]-res3_km[i,-1]
    tmp <- rbind(tmp1,tmp2,tmp3)
    rownames(tmp) <- paste0(rownames(res1_km)[i],"_",c("GSE75748","GSE81861","cellbench"))
    tmp_median <- apply(tmp,2,median)
    res_km <- rbind(res_km,tmp)
    res_km_median <- rbind(res_km_median,tmp_median)
}

for(i in 3:5){
    tmp1 <- res1_km[i,-1]-res1_km[i,1]
    tmp2 <- res2_km[i,-1]-res2_km[i,1]
    tmp3 <- res3_km[i,-1]-res3_km[i,1]
    tmp <- rbind(tmp1,tmp2,tmp3)
    rownames(tmp) <- paste0(rownames(res1_km)[i],"_",c("GSE75748","GSE81861","cellbench"))
    tmp_median <- apply(tmp,2,median)
    res_km <- rbind(res_km,tmp)
    res_km_median <- rbind(res_km_median,tmp_median)
}

res_km_median <- as.data.frame(t(res_km_median))
rownames(res_km_median) <- colnames(res1_km)[-1]
colnames(res_km_median) <- c("Hacc","Hpur","ARI","medianSil","NMI")

res_km_median$method <- rownames(res_km_median)
res_km_median_melt <- data.table::melt(res_km_median,id.vars=c("method"))
colnames(res_km_median_melt)[2:3] <- c("Metrics","Metrics_Difference")
head(res_km_median_melt)

res_km$Metrics_Datasets <- rownames(res_km)
res_km$Metrics <- sapply(rownames(res_km),function(x) strsplit(x,"\\_")[[1]][1] )
res_km_melt <- data.table::melt(res_km,id.vars=c("Metrics_Datasets","Metrics"))
colnames(res_km_melt)[3:4] <- c("method","Metrics_Difference")
head(res_km_melt)
```

heatmap(median metrics)
```{r}
rank1 <- apply(res_km_median[,-ncol(res_km_median)],2,rank)
rank1 <- apply(rank1,1,median)
rank1 <- names(rank1[order(rank1)])
res_km_median_melt$Method <- factor(res_km_median_melt$method,levels=unique(res_km_median_melt$method[order(res_km_median_melt$method)]))
res_km_median_melt$method <- factor(res_km_median_melt$method,levels=rank1)
res_km_median_melt <- res_km_median_melt[res_km_median_melt$method!="New",]

tmp_hm <- res_km_median_melt[,]
tmp_hm[tmp_hm$Metrics_Difference<(-0.01),]$Metrics_Difference <- -0.01
tmp_hm[tmp_hm$Metrics_Difference>(0.01),]$Metrics_Difference <- 0.01
tmp_hm <- tmp_hm[tmp_hm$Metrics!="medianSil",]
clu_4metric_km_hm <- ggplot(tmp_hm,aes(method,Metrics,fill=Metrics_Difference))+geom_tile()+theme(axis.text.x=element_text(angle=90))+scale_fill_gradient2(low="blue",mid="white",high="red",midpoint=0)+ggtitle("K-means Clustering Evaluation Metrics(Values of Difference)")+xlab("Method")+theme(legend.title=element_text(size=15),legend.text=element_text(size=14),axis.text=element_text(size=15),axis.title=element_text(size=15),plot.title=element_text(size=15))#+geom_vline(xintercept=neg_md+0.5,linetype="dashed")
clu_4metric_km_hm
```

# cell cycle dynamics
## GSE64016
### preprocess
```{r message=FALSE, warning=FALSE}
##raw
raw_dat <- fread("rawcount_txt/GSE64016_genebycell.txt",sep="\t",header=TRUE,stringsAsFactors=FALSE)
raw_dat <- as.matrix(raw_dat,rownames=1)
##imputed
MAGIC_dat <- fread("MAGIC/GSE64016_genebycell.tsv",sep="\t",header=TRUE,stringsAsFactors=FALSE)
MAGIC_dat <- as.matrix(MAGIC_dat,rownames=1)
alra_dat <- fread("alra/GSE64016_genebycell_alra.txt",sep="\t",stringsAsFactors=FALSE)
alra_dat <- as.matrix(alra_dat,rownames=1)
dca_dat <- fread("dca/GSE64016/mean.tsv",sep="\t",header=TRUE,stringsAsFactors=FALSE)
dca_dat <- as.matrix(dca_dat,rownames=1)
ks_dat <- fread("knn_smoothing/GSE64016_genebycell_ks.txt",sep="\t",header=TRUE,stringsAsFactors=FALSE)
ks_dat <- as.matrix(ks_dat,rownames=1)
MAGIClog_dat <- fread("MAGIC_log/GSE64016_genebycell.tsv",sep="\t",header=TRUE,stringsAsFactors=FALSE)
MAGIClog_dat <- as.matrix(MAGIClog_dat,rownames=1)
AutoClass_dat <- fread("AutoClass/GSE64016_genebycell_AC.txt",sep="\t",header=TRUE,stringsAsFactors=FALSE)
AutoClass_dat <- as.matrix(AutoClass_dat,rownames=1)
ccImpute_dat <- fread("ccImpute/GSE64016_genebycell_cc.txt",sep="\t",stringsAsFactors=FALSE)
ccImpute_dat <- as.matrix(ccImpute_dat,rownames=1)
Iimpute_dat <- fread("I_Impute/GSE64016_genebycell_CImp.txt",sep="\t",header=TRUE,stringsAsFactors=FALSE)
Iimpute_dat <- as.matrix(Iimpute_dat,rownames=1)
scRMDnorm_dat <- fread("scRMD/GSE64016_genebycell_scRMDlognorm.txt",sep="\t",stringsAsFactors=FALSE)
scRMDnorm_dat <- as.matrix(scRMDnorm_dat,rownames=1)
Bf_dat <- fread("Bfimpute/GSE64016_genebycell_Bfimpute_specc.txt",sep="\t",stringsAsFactors=FALSE)
Bf_dat <- as.matrix(Bf_dat,rownames=1)
afMF_dat <- fread("afMF/imputed_data/GSE64016_sigma3_0_convergence_True.txt",sep="\t",header=TRUE,stringsAsFactors=FALSE)
afMF_dat <- as.matrix(afMF_dat,rownames=1)

metadat <- sapply(colnames(raw_dat),function(x) strsplit(x,"\\_")[[1]][1])
metadat <- data.frame(cell=colnames(raw_dat),celltype=metadat)
```

normalized data
```{r}
raw_dat <- CreateSeuratObject(counts=raw_dat,meta.data=metadat,min.cells=0,min.features=0)
raw_dat <- NormalizeData(raw_dat)
raw_dat <- as.matrix(raw_dat@assays$RNA@data)

dca_dat <- log2(dca_dat+min(dca_dat[dca_dat!=0]))

ks_dat <- CreateSeuratObject(counts=ks_dat,meta.data=metadat,min.cells=0,min.features=0)
ks_dat <- NormalizeData(ks_dat)
ks_dat <- as.matrix(ks_dat@assays$RNA@data)

Iimpute_dat <- log2(Iimpute_dat)
```

combine normalized data
```{r}
dat <- list(raw=raw_dat,ALRA=alra_dat,AutoClass=AutoClass_dat,Bfimpute=Bf_dat,ccImpute=ccImpute_dat,DCA=dca_dat,I_Impute=Iimpute_dat,kNN_smoothing=ks_dat,MAGIC=MAGIC_dat,MAGIC_log=MAGIClog_dat,scRMD=scRMDnorm_dat,afMF=afMF_dat)
rm(raw_dat)
rm(alra_dat)
rm(AutoClass_dat)
rm(Bf_dat)
rm(ccImpute_dat)
rm(Iimpute_dat)
rm(ks_dat)
rm(dca_dat)
rm(MAGIC_dat)
rm(MAGIClog_dat)
rm(scRMDnorm_dat)
rm(afMF_dat)

colnames(dat$ALRA) <- colnames(dat$raw)

#keep cells with cell cycle label
metadat <- metadat[metadat$celltype!="H1",]
metadat[metadat$celltype=="G2",]$celltype <- "G2M"
metadat$Phase <- factor(metadat$celltype,levels=c("G1","S","G2M"))
dat <- lapply(dat,function(x) x[,metadat$cell] )
```

### clustering for cells:cell cycle scores with Seurat
```{r}
sgenes <- cc.genes$s.genes
g2mgenes <- cc.genes$g2m.genes
AllccGenes <- unique(c(sgenes,g2mgenes))

getSeuratCC <- function(x){
    tmp <- CreateSeuratObject(x,min.cells=0,min.features=0)
    tmp <- ScaleData(tmp,features=rownames(tmp))
    tmp <- CellCycleScoring(tmp,s.features=sgenes,g2m.features=g2mgenes,set.ident=TRUE,seed=42)
    tmp <- RunPCA(tmp,features=AllccGenes,verbose=FALSE)
    tmp <- RunUMAP(tmp,verbose=FALSE,dims=1:50)
    return(tmp)
}

dat2 <- lapply(dat,getSeuratCC)

getTruePhase <- function(x){
    x$TruePhase <- metadat$celltype
    return(x)
}
dat2 <- lapply(dat2,getTruePhase)
```

#### eval acc and f1
```{r}
evalPhaseAcc <- function(x,metad){
    acc <- sum(x$Phase==metad$celltype)/nrow(metad)
    return(acc)
}

evalCellCycleP <- function(x,metad,phase){
    phaseCol <- paste0(phase,".Score")
    res1 <- x@meta.data[metad$celltype==phase,phaseCol]
    res2 <- x@meta.data[metad$celltype!=phase,phaseCol]
    res <- wilcox.test(res1,res2,alternative="greater")$p.value
    return(res)
}

evalPhasemedianF1 <- function(x,metad){
    tmp <- confusionMatrix(factor(x$Phase,levels=levels(metad$Phase)),metad$Phase,mode="everything")
    tmp <- median(tmp$byClass[,"F1"],na.rm=TRUE)
    return(tmp)
}

evalPhaseF1 <- function(x,metad){
    tmp <- confusionMatrix(factor(x$Phase,levels=levels(metad$Phase)),metad$Phase,mode="everything")
    tmp <- tmp$byClass[,"F1"]
    return(tmp)
}

```

```{r}
##accracy
res_acc1 <- as.data.frame(lapply(dat2,evalPhaseAcc,metadat))
res_acc1 <- as.data.frame(t(res_acc1))
res_acc1

#F1 median
res_mf1 <- as.data.frame(lapply(dat2,evalPhasemedianF1,metadat))
res_mf1 <- as.data.frame(t(res_mf1))
res_mf1

#cell cycle p
res_ccp1 <- as.data.frame(lapply(dat2,evalCellCycleP,metadat,"S"))
res_ccp2 <- as.data.frame(lapply(dat2,evalCellCycleP,metadat,"G2M"))
res_ccp <- rbind(res_ccp1,res_ccp2)
res_ccp <- as.data.frame(t(res_ccp))
res_ccp

#F1 all
res_f1 <- as.data.frame(lapply(dat2,evalPhaseF1,metadat))
res_f1
```

barplot
Acc
```{r}
res_acc1$methods <- rownames(res_acc1)
rank1 <- res_acc1[order(res_acc1$V1),]$methods
res_acc1$Method <- factor(res_acc1$methods,levels=c("raw",res_acc1$methods[-1][order(res_acc1$methods[-1])]))
res_acc1$method <- factor(res_acc1$methods,levels=rank1)

res_acc1$Type <- "Deep learning based"
res_acc1[res_acc1$method%in%c("ADimpute","ccImpute","I_Impute","kNN_smoothing","MAGIC","MAGIC_log","RESCUE"),]$Type <- "Model or smoothing based"
res_acc1[res_acc1$method%in%c("ALRA","Bfimpute","afMF","scRMDraw","scRMDnorm"),]$Type <- "Matrix based"
res_acc1[res_acc1$method%in%c("raw"),]$Type <- "Raw"
res_acc1$Type <- factor(res_acc1$Type,levels=c("Matrix based","Model or smoothing based","Deep learning based","Raw"))

colnames(res_acc1)[1] <- "Accuracy"

tmp_bar <- res_acc1[,]

CCD_acc_bp <- ggbarplot(tmp_bar,x="method",y="Accuracy",fill="Type",position=position_dodge(0.9),xlab="Method",ylab="Accuracy,%")+ggtitle("Predicted Cell Cycle Accuracy")+geom_hline(yintercept=res_acc1[res_acc1$methods=="raw","Accuracy"],linetype="dashed")+coord_cartesian(ylim=c(0.25,0.65))+theme(plot.title=element_text(size=16),legend.title=element_text(size=15),legend.text=element_text(size=14),axis.title=element_text(size=18),axis.text.x=element_text(angle=90,size=15),axis.text.y=element_text(size=15),legend.position="right")+scale_fill_brewer(palette="Set2")
CCD_acc_bp
```

F1
```{r}
res_mf1$methods <- rownames(res_mf1)
rank1 <- res_mf1[order(res_mf1$V1),]$methods
res_mf1$Method <- factor(res_mf1$methods,levels=c("raw",res_mf1$methods[-1][order(res_mf1$methods[-1])]))
res_mf1$method <- factor(res_mf1$methods,levels=rank1)
colnames(res_mf1)[1] <- "F1"

res_mf1$Type <- "Deep learning based"
res_mf1[res_mf1$method%in%c("ADimpute","ccImpute","I_Impute","kNN_smoothing","MAGIC","MAGIC_log","RESCUE"),]$Type <- "Model or smoothing based"
res_mf1[res_mf1$method%in%c("ALRA","Bfimpute","afMF","scRMDraw","scRMDnorm"),]$Type <- "Matrix based"
res_mf1[res_mf1$method%in%c("raw"),]$Type <- "Raw"
res_mf1$Type <- factor(res_mf1$Type,levels=c("Matrix based","Model or smoothing based","Deep learning based","Raw"))

tmp_bar <- res_mf1[,]

CCD_f1_bp <- ggbarplot(tmp_bar,x="method",y="F1",fill="Type",position=position_dodge(0.9),xlab="Method",ylab="F1 score")+ggtitle("Predicted Cell Cycle F1 score(Median)")+geom_hline(yintercept=res_mf1[res_mf1$methods=="raw","F1"],linetype="dashed")+coord_cartesian(ylim=c(0.35,0.8))+theme(plot.title=element_text(size=16),legend.title=element_text(size=15),legend.text=element_text(size=14),axis.title=element_text(size=18),axis.text.x=element_text(angle=90,size=15),axis.text.y=element_text(size=15),legend.position="right")+scale_fill_brewer(palette="Set2")
CCD_f1_bp
```

Pvalue
```{r}
colnames(res_ccp) <- c("S vs Others","G2M vs Others")
res_ccp$`S vs Others` <- log10(res_ccp$`S vs Others`)*(-1)
res_ccp$`G2M vs Others` <- log10(res_ccp$`G2M vs Others`)*(-1)
res_ccp$methods <- rownames(res_ccp)
res_ccp_melt <- data.table::melt(res_ccp,id.vars=c("methods"),variable.name="Cell_Cycle",value.name="log10P")
head(res_ccp_melt)
```

barplot
```{r}
res_ccp_melt$Method <- factor(res_ccp_melt$methods,levels=c("raw",res_ccp$methods[-1][order(res_ccp$methods[-1])]))

tmp_bar <- res_ccp_melt[,]
CCD_compare_bp <- ggbarplot(tmp_bar,x="Cell_Cycle",y="log10P",fill="Method",palette=selected_palette2,position=position_dodge(0.9),xlab="Cell Cycle Comparison",ylab="-log10(P)")+ggtitle("Comparison of predicted cell cycle score between known cell cycle status")+theme(legend.title=element_text(size=15),legend.text=element_text(size=14),axis.text=element_text(size=15),axis.title=element_text(size=15),plot.title=element_text(size=15))
CCD_compare_bp
```

#### pca and umap
pca(not clear; removed)
```{r}
for(x in names(dat2)){
    print(DimPlot(dat2[[x]],group.by="TruePhase",reduction="pca",pt.size=5)+ggtitle(x))
}
```

umap true phase
```{r fig.height=8, fig.width=12}
p_list <- list()

rank_name <- c("raw",names(dat2)[-1][order(names(dat2)[-1])])
for(x in rank_name){
    p_list[[x]] <- DimPlot(dat2[[x]],group.by="TruePhase",reduction="umap",pt.size=5)+ggtitle(x)+scale_color_brewer(palette="Set2")
}

CCD_umap_true <- ggarrange(plotlist=p_list[c("raw","MAGIC","AutoClass","DCA","afMF","ALRA","ccImpute","I_Impute","Bfimpute","scRMD","kNN_smoothing","MAGIC_log")],ncol=4,nrow=3)
CCD_umap_true
```


# combine plot
## main
```{r fig.height=7.4, fig.width=11}
#h12.3 w11 for 3
p1 <- ggarrange(clu_4metric_lv_hm,clu_4metric_km_hm,CCD_compare_bp,ncol=3,labels="AUTO")
p2 <- ggarrange(CCD_umap_true,labels=c("D"))

ggarrange(p1,p2,nrow=2,heights=c(1,2))
```

## supp1
```{r fig.height=2.5,fig.width=8}
ggarrange(CCD_acc_bp,CCD_f1_bp,ncol=2,nrow=1,labels="AUTO")
```

## supp2
```{r fig.height=5, fig.width=11}
ggarrange(difftraj_umap2_raw,difftraj_umap2_afMF,difftraj_umap2_alra,difftraj_umap2_ac,difftraj_umap2_magic,ncol=3,nrow=2,labels="AUTO")
```



